### 2016-02-27 docs/paper/response.txt
reviewer 2 minor comments
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2241 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

===================================================

Reviewer 1

The first question in the second paragraph appears to be more
for direct information to the reviewer than request for inclusion 
in the paper. If that isn't so, we are happy to address this
there (though it seems a bit off the software issue). For HST
there is a signoff on updates by instrument teams, and there is
a committee responsible for the updates, but I don't think I would
go as far as to call it a CCB. As far as contemporaneous calibration
data, these are invariably associated with the proposal and thus
the calibration is done through associating the data with the science
data in the pipeline, and these files generally do not appear
in the CDRS (or the older CDBS). Pipelines are fed tables of the
files that belong in an association so that they know which file
to use for such a purpose.

The issue of tracability is handled by the fact that all data products
are tagged with the version of CRDS rules in their header
(the pipeline context) along with relevant software versions.
Given that, the exact same numbers should be extracted if the 
same context is given. That was one major goal of the system was 
to be able to replicate previous results (file recommendations
and otherwise), which the older system could not do. We have 
added a sentence assuring that this functionality is also repeatable
for the given context. E.g., one knows which rules file the value
was extracted from and that provides all the information one 
needs.

The comment about limiting the downloads needed for areas with
limited bandwidth is already a feature of the system. The client
library only downloads files that a pipeline invocation needs, and
these downloads will be cached for the event other datasets 
require them. No other files are automatically downloaded when it
is used in this manner. We have added a sentence to make that clear.

Section 2 comments
------------------

Paper section 3 item 1:
The main point being made was that one team may submit a change
to be tested (which could take a while), followed by another team
that has a  more urgent need for an update and tesing, yet must
await the previous test to complete. The teams operate asynchronously
so one test system isn't necessarily sufficient. But the criticism 
is a good one, and given this isn't a very important point in the
paper, we have shortened this issue to say essentially that the
test system didn't address all test needs.

page 6, substitutions.
The wording suggestions are indeed much better and have been 
made.

page 6, last paragraph
Again the suggested text is much better and has been used.

Section 3 comments
------------------

Figures 1-6 have been boxed as requested (we do agree that this
is a signficant improvement)

Figures 7-12 xxxx

Section 4 comments
------------------

These have all been fixed (we hope :-) We thank the reviewer for
taking the time to detail these. After viewing the same or similar
text repeatedly it is quite hard to look as closely as one would
wish, so this is quite helpful.

****** To do

1) box text file figures
2) improve screen shots

===================================================

Reviewer 2

Section 2
We have reworded to be more explicit that all possible combinations of parameters
must be represented as distinct rows.

Section 3 point 2: "Could an 'is valid' boolean attribute in the database
help out?"
Not really. The file itself may be perfectly good, it is just associated with
the wrong value of parameters. Possbily if coupled with a resubmission with the
right values, but note that may mean many entries have to be so marked, and 
replicated. But as implemented, it didn't have that option.

Section 5: first sentence.
Fixed.
Section 5 submission typo
Fixed
Section 5 point 3
todo****xxxx

Section 6.3 paragraph 2
reworded
Section 6.3 par 4
Made it clear that it is the version of the calibration pipeline that is 
being referred to.
Section 6.4.1 Relational Expressions typo
Fixed
Section 6.4.1 next to last par
Joined two sentences properly.
Section 6.5 par 3
Fixed
Section 7 par 2 punctuation
Fixed
Section 7.2 par 2 begains 
Fixed