### 2016-03-14 docs/paper/response.txt
updated all text except for pseudo-mode clarification
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2248 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

===================================================

Reviewer 1

The first question in the second paragraph appears to be more
for direct information to the reviewer than request for inclusion 
in the paper. If that isn't so, we are happy to address this
there (though it seems a bit off the software issue). For HST
there is a signoff on updates by instrument teams, and there is
a committee responsible for the updates, but I don't think I would
go as far as to call it a CCB. As far as contemporaneous calibration
data, these are invariably associated with the proposal and thus
the calibration is done through associating the data with the science
data in the pipeline, and these files generally do not appear
in the CDRS (or the older CDBS). Pipelines are fed tables of the
files that belong in an association so that they know which file
to use for such a purpose.

The issue of tracability is handled by the fact that all data products
are tagged with the version of CRDS rules in their header
(the pipeline context) along with relevant software versions.
Given that, the exact same numbers should be extracted if the 
same context is given. That was one major goal of the system was 
to be able to replicate previous results (file recommendations
and otherwise), which the older system could not do. We have 
added a sentence assuring that this functionality is also repeatable
for the given context. E.g., one knows which rules file the value
was extracted from and that provides all the information one 
needs.

The comment about limiting the downloads needed for areas with
limited bandwidth is already a feature of the system. The client
library only downloads files that a pipeline invocation needs, and
these downloads will be cached for the event other datasets 
require them. No other files are automatically downloaded when it
is used in this manner. We have added a sentence to make that clear.

Section 2 comments
------------------

Paper section 3 item 1:
The main point being made was that one team may submit a change
to be tested (which could take a while), followed by another team
that has a  more urgent need for an update and tesing, yet must
await the previous test to complete. The teams operate asynchronously
so one test system isn't necessarily sufficient. But the criticism 
is a good one, and given this isn't a very important point in the
paper, we have shortened this issue to say essentially that the
test system didn't address all test needs.

page 6, substitutions.
The wording suggestions are indeed much better and have been 
made.

page 6, last paragraph
Again the suggested text is much better and has been used.

Section 3 comments
------------------

Figures 1-6 have been boxed as requested (we do agree that this
is a signficant improvement)

Figures 7-12 xxxx

Section 4 comments
------------------

These have all been fixed (we hope :-) We thank the reviewer for
taking the time to detail these. After viewing the same or similar
text repeatedly it is quite hard to look as closely as one would
wish, so this is quite helpful.

****** To do

1) box text file figures
2) improve screen shots

===================================================

Reviewer 2

General comment regarding databases and comparisions to other observatories.

A section at the end has been added that briefly describes what Spitzer, 
Chandra and Gemini use and notes some of the restrictions entailed in their
approaches (Chandra's is most similar in approach but simpler).

As far as the implications regarding databases go, we have not expanded on that.
This is mainly because some issues are arguably self evident (distributing
software that requires databases certainly does impose a significantly larger 
burden on those installing and maintaining such a system; we doubt that is 
controversial).

We also think it is nearly self evident that SQL does not always represent 
rules in a clear or easy to understand way (and many astronomers, the ones 
that ultimately must provide those rules, are not experts in SQL, particularly
complex SQL). We aren't saying it is impossible to use only SQL, just not 
ideal. Indeed, Gemini's experience found it to be too constricting (as we did).
Chandra doesn't use it at all, and it appears Spitzer's needs are not as complex.
This is probably aided by the fact that they generally process the data in 
large runs, i.e,. their updates are relatively infrequent as we understand 
it.  Besides, they sometimes have special processing after the query (e.g., 
interpolation between two reference files), so it isn't a pure database
solution either. Accomodating new needs requires new software changes.
To be fair, so does ours if a new kind of rule is required, but that
is much less likely, and much more likely to be in adding a new module
rather than changing existing code.

As to the focus on STScI needs. It would be nice if we had the resources 
to develop a system with a very broad reach and goals initially (well, perhaps,
many very ambitious projects die for being too ambitious!). But with the
reality of funding oversight, it's very difficult to do that. NASA gets 
quite nervous when they see work being done that isn't specifically done
for their mission. We do what we can to keep the project more broadly 
useful within these constraints. That's even true when writing papers
(i.e., can we justify spending much time doing a thorough survey of all
systems out there?).

Section 2
We have reworded to be more explicit that all possible combinations of parameters
must be represented as distinct rows.

Section 3 point 2: "Could an 'is valid' boolean attribute in the database
help out?"
Not really. The file itself may be perfectly good, it is just associated with
the wrong value of parameters. Possbily if coupled with a resubmission with the
right values, but note that may mean many entries have to be so marked, and 
replicated. But as implemented, it didn't have that option. If it had been 
designed from the beginning to handle that, yes, it could have dealt with bad
submissions that way.

Section 5: first sentence.
Fixed.
Section 5 submission typo
Fixed
Section 5 point 3
todo****xxxx

Section 6.3 paragraph 2
reworded
Section 6.3 par 4
Made it clear that it is the version of the calibration pipeline that is 
being referred to.
Section 6.4.1 Relational Expressions typo
Fixed
Section 6.4.1 next to last par
Joined two sentences properly.
Section 6.5 par 3
Fixed
Section 7 par 2 punctuation
Fixed
Section 7.2 par 2 begains 
Fixed